---
title: "Introduction to Nutrilite: Validation with Simulated Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Nutrilite: Validation with Simulated Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## 1. Introduction

The `Nutrilite` package provides a robust framework for estimating nutrient uptake rates in stream ecosystems. It uses Bayesian optimization (`mlr3mbo`) to fit advection-dispersion equations, supporting transient storage and different uptake kinetics.

To demonstrate the accuracy of our two-step fitting workflow, we use a **simulated dataset** with known "ground truth" parameters. The true parameters for this First-Order with Transient Storage (`FO_TS`) model are:
* **Hydraulic**: $D = 0.5$, $U = 1.5$, $\alpha = 0.05$, $A_s/A = 2.0$
* **Biogeochemical**: $K_N = 0.05$
* **Background**: $Cl_{bg} = 6.0$, $N_{bg} = 3.0$

## 2. Setup and Data Preparation

First, we load the package and our simulated breakthrough curve (BTC) data for both the conservative tracer (Chloride) and the reactive tracer (Nitrogen).

```{r setup, eval=FALSE}
library(Nutrilite)

# Use system.file() to locate the raw CSV files shipped with the package
cl_file <- system.file("extdata", "simu_fo_ts_cl.csv", package = "Nutrilite")
n_file <- system.file("extdata", "simu_fo_ts_n.csv", package = "Nutrilite")

# Read the simulated dataset
df_cl <- read.csv(cl_file)
df_n <- read.csv(n_file)

time_vec <- df_cl$time
cl_conc <- df_cl$cl_obs
n_conc <- df_n$n_obs

# Physical site and injection parameters
distance_m <- 40
cross_area_m2 <- 1.0
release_zone <- 0.4
volume <- cross_area_m2 * release_zone

mass_cl <- 1500 * volume  # 600 mg
mass_n <- 2500 * volume   # 1000 ug
```

## 3. Step 1: Fitting the Conservative Tracer

We fit the conservative tracer first to estimate the stream's physical transport properties. We optimize $D$, $U$, $\alpha$, $A_s/A$, and the background concentration.

```{r fit-cl, eval=FALSE}
# Note: n_evals is set to 70 for vignette speed. For real data, 80-150 is recommended.
fit_cl <- nutrilite(
  time = time_vec, 
  conc = cl_conc, 
  distance = distance_m, 
  injected_mass = mass_cl, 
  cross_area = cross_area_m2,
  tracer_type = "conservative", 
  model_type = "FO_TS", 
  bounds = list(
    D = c(0.1, 1),        
    U = c(1.35, 1.65),        
    Alpha = c(0.01, 0.1),   
    AsA_coeff = c(0.5, 3.5),
    bg_conc = c(5.5, 6.5)    
  ),
  n_evals = 70
)

summary(fit_cl)
# The estimated parameters should be very close to the true values:
# D ~ 0.5, U ~ 1.5, Alpha ~ 0.05, AsA_coeff ~ 2.0
```

## 4. Step 2: Fitting the Reactive Tracer

Now we extract the optimized hydraulic parameters and pass them as `fixed_params` to isolate and model the reactive nitrogen dynamics.

```{r fit-n, eval=FALSE}
hydro_params <- fit_cl$best_params

fit_n <- nutrilite(
  time = time_vec, 
  conc = n_conc, 
  distance = distance_m, 
  injected_mass = mass_n, 
  cross_area = cross_area_m2,
  tracer_type = "reactive", 
  model_type = "FO_TS", 
  fixed_params = list(
    D = hydro_params$D, 
    U = hydro_params$U, 
    Alpha = hydro_params$Alpha, 
    AsA_coeff = hydro_params$AsA_coeff
  ),
  bounds = list(
    K_N = c(0.01, 0.1),     
    bg_conc = c(2.5, 3.5)   
  ),
  n_evals = 70
)

summary(fit_n)
# The estimated uptake rate K_N should closely approximate the true value of 0.05.
```

## 5. Visualizing the Fit

The package includes S3 plot methods to easily visualize how well the fitted models match the observed breakthrough curves.

```{r plot-fits, eval=FALSE}
par(mfrow=c(1,2)) 
plot(fit_cl, main="Conservative Tracer Fit (Cl)")
plot(fit_n, main="Reactive Tracer Fit (N)")
par(mfrow=c(1,1))
```

The resulting plots will display the observed data points alongside the smooth predicted curves generated by the optimized parameters, confirming the accuracy of the `Nutrilite` inference engine.